ST563 Final Project

### 2021-03-11 ST563-FP.Rmd
added code for MLR model
fixed issue with kable outputs

# Load necessary libraries
library(tidyverse) # so many things
library(knitr) # making nice tables
library(caret) # training models & CV
library(leaps) # subset selection
library(corrplot) # visualize correlation
library(rmarkdown) # for output documents
library(rpart) # fitting classification tree
library(rpart.plot) # visualization of classification tree
library(randomForest) # fitting random forest model
library(kernlab) # fitting SVM
library(class) # fitting kNN model
library(glmnet) # fitting LASSO 
library(gbm) # fitting the boosted tree model

# Read in data set
wine <- read.csv2(file = "winequality-red.csv", header = TRUE, dec = ".",
                  colClasses = c(rep("double", 7), rep("double", 4), "integer"))
```

@@

```{r methods}
cl <- c("Classification Tree", "Random Forest", "Support Vector Machine (SVM)",
        "KNN Classification", "Logistic Regression")
reg <- c("Regression Tree", "Boosted Tree", "Multiple Linear Regression (MLR)", 
          "KNN Regression", "LASSO Regression")
kable(data.frame("Classification" = cl, "Regression" = reg), caption = "ML Methods")
```

@@

## Variable Selection
After running an initial random forest model on the entire dataset with selected predictors, we looked at the variable importance of the predictors to decide which variables we would use in our final model. Variable importance, in the context on machine learning, refers to how much a given model "uses" that variable to make accurate predictions. In other words, the more a model relies on a variable to make predictions, the more important it is for the model. 
```{r varImp, echo = FALSE}
# Write preliminary model  
model.class.pre <- class ~ .

# Fit preliminary random forest on entire data set
rf.class.pre <- randomForest(model.class.pre, data=wine, ntree=500, mtry=3, importance=TRUE)

@@

# Plot variable importance
vimp.class <- ggplot(VIrf.pre.class.sort[-12,],aes(x=X,y=MeanDecreaseAccuracy))+
              geom_bar(stat = 'identity', position = 'dodge', fill="lightblue")+
              theme(axis.text.x = element_text(angle = 60, hjust = 1))+
              ylab("Mean Decrease in Accuracy") + xlab("Predictor")+
              ggtitle("Variable Importance Plot")
vimp.class
```

```{r}
# Create model and get summary output
lm.sub <- regsubsets(quality ~ ., data = wine, nvmax = 10)
smre <- lm.sub %>% summary()

# Plot the Cp, BIC, and R^2 results in one plot
par(mfrow = c(2,2))
plot(smre$cp, xlab = "# of Preds", ylab = "Cp")
points(which.min(smre$cp), smre$cp[which.min(smre$cp)], pch = 8, col = "#CC0000")
plot(smre$bic, xlab = "# of Preds", ylab = "BIC")
points(which.min(smre$bic), smre$bic[which.min(smre$bic)], pch = 8, col = "#CC0000")
plot(smre$adjr2, xlab = "# of Preds", ylab = "Adjusted R^2")
points(which.max(smre$adjr2), smre$adjr2[which.max(smre$adjr2)], pch = 8, col = "#CC0000")

@@

```{r form}
form.reg <- quality ~ volatile.acidity + chlorides + 
            total.sulfur.dioxide + pH + sulphates + alcohol
form.class <- class ~ alcohol + sulphates + volatile.acidity + density + total.sulfur.dioxide
```

@@

```{r tree, echo=FALSE}
#### REGRESSION ####
# Fit basic regression tree model
tree.reg <- train(form.reg, data = trn, method = "rpart",
                  trControl = trctrl)

# Obtain RMSE and results
tree.reg.pred <- predict(tree.reg, newdata = tst)
tree.reg.rmse <- sqrt(mean((tree.reg.pred - tst$quality)^2))

#### HOTEL 2 ####
# Fit classification tree model
tree.class <- train(form.class, data = trn, method = "rpart",
                    trControl = trctrl)

@@

# Print out classification tree visualization
#tree.class.out <- rpart(form.class, data = trn, cp = 0.01971326)
#rpart.plot(tree.class.out, main="Classification Tree")

# Print out confusion matrix
#list("RMSE" = tree.reg.rmse,
#     "Class Results" = tree.class.res)
```

@@

```{r rf, echo=FALSE}
#### REGRESSION ####
# Fit the regression random forest model 
rf.reg <- train(form.reg, data = trn, method = "rf",
                trControl = trctrl, importance = TRUE)

@@

To fit the LASSO regression model, we used the `glmnet` package in R.
```{r}
#### REGRESSION ####
# Fit LASSO to the data
lasso.reg <- train(form.reg, data = trn, method = "glmnet", trControl = trctrl, 
                   tuneGrid = expand.grid(.alpha=1, .lambda=0.00175))

@@

```{r}
#### REGRESSION ####
# Fit the regression boosted tree model 
boost.reg <- train(form.reg, data = trn, method = "gbm",
                trControl = trctrl, verbose = FALSE)

@@

@@

>>>>

# Load necessary libraries
library(tidyverse) # so many things
library(knitr) # making nice tables
library(caret) # training models & CV
library(leaps) # subset selection
library(corrplot) # visualize correlation
library(rmarkdown) # for output documents
library(rpart) # fitting classification tree
library(rpart.plot) # visualization of classification tree
library(randomForest) # fitting random forest model
library(kernlab) # fitting SVM
library(class) # fitting kNN model
library(glmnet) # fitting LASSO 
library(gbm) # fitting the boosted tree model
library(kableExtra) # making bolded headers for tables

# Read in data set
wine <- read.csv2(file = "winequality-red.csv", header = TRUE, dec = ".",
                  colClasses = c(rep("double", 7), rep("double", 4), "integer"))

# Set seed for reproducibility
set.seed(702)
```

@@

```{r methods}
cl <- c("Classification Tree", "Random Forest", "Support Vector Machine (SVM)",
        "KNN Classification", "Logistic Regression")
reg <- c("Boosted Tree", "Random Forest", "Multiple Linear Regression (MLR)", 
          "KNN Regression", "LASSO Regression")
kable(data.frame("Classification" = cl, "Regression" = reg), caption = "ML Methods") %>% 
  row_spec(0, bold=TRUE) %>% collapse_rows(columns = 1, latex_hline = "none") %>% 
  kable_styling(latex_options = "hold_position")
```

@@

## Variable Selection
Since we have two different tasks that we are trying to complete (classification and regression), we went about selecting the variables for each task differently. For the classification task, we ran an initial random forest model on the entire dataset with selected predictors. We then looked at the variable importance of the predictors to decide which variables we would use in our final models. Variable importance, within the context of machine learning, refers to how much a given model "uses" that variable to make accurate predictions. In other words, the more a model relies on a variable to make predictions, the more important it is for the model.

```{r class.select, fig.align="center", out.width='4in'}
# Write preliminary model  
model.class.pre <- class ~ .

# Fit preliminary random forest on entire data set
rf.class.pre <- randomForest(model.class.pre, data = wine[,-12], 
                             ntree = 500, mtry = 3, importance = TRUE)

@@

# Plot variable importance
vimp.class <- ggplot(VIrf.pre.class.sort, aes(x = X, y = MeanDecreaseAccuracy)) +
                geom_bar(stat = 'identity', position = 'dodge', fill="lightblue") +
                theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
                ylab("Mean Decrease in Accuracy") + xlab("Predictor") +
                ggtitle("Variable Importance") + 
                geom_hline(yintercept = .005, lty = 2, colour = "darkblue")
vimp.class
```

We decided to choose the variables with a variable importance over .005, leaving us with `alcohol`, `sulphates`, `volatile.acidity`, `density`, and `total.sulfur.dioxide`. The fact that density proved to be an important variable for classification was relatively surprising. One would think that the wine quality would be determined based on its attributes that are detectable by the human senses. For this same reason, alcohol, acidity, and sulphates made sense in terms of variable importance. Sulphates are essential for the preservation of the wine, an overly- or underly-acidic wine may be considered poor, and the alcohol content has a connection to the fermentation process, where the must (aka the grape juice before it undergoes fermentation) actually becomes wine.

When looking at the problem in terms of a regression context, we used best subsets selection for feature selection. Using the `leaps` library and `regsubsets` in R, we are able to compare all of the possible models from the given set of predictors. Through best subset selection, we can get the set of predictors for a model with any given amount of predictors. We looked at three different model criteria to compare models of different amounts of predictor variables: Mallow's Cp, BIC, and $R^2_{\text{adj}}$.

```{r reg.select}
# Create model and get summary output
lm.sub <- regsubsets(quality ~ ., data = wine[,-13], nvmax = 11)
smre <- lm.sub %>% summary()

# Plot the Cp, BIC, R^2, RSS results in one plot
par(mfrow = c(2,2))
plot(smre$cp, xlab = "# of Preds", ylab = "Cp")
points(which.min(smre$cp), smre$cp[which.min(smre$cp)], pch = 8, col = "#CC0000")
plot(smre$bic, xlab = "# of Preds", ylab = "BIC")
points(which.min(smre$bic), smre$bic[which.min(smre$bic)], pch = 8, col = "#CC0000")
plot(smre$adjr2, xlab = "# of Preds", ylab = "Adjusted R^2")
points(which.max(smre$adjr2), smre$adjr2[which.max(smre$adjr2)], pch = 8, col = "#CC0000")
plot(smre$rss, xlab = "# of Preds", ylab = "Training RSS")
points(which.min(smre$rss), smre$rss[which.min(smre$rss)], pch = 8, col = "#CC0000")

@@

Although each of the three model criteria suggested a different number of predictor variables of a different number, they were all within the same region (between 6 and 8 predictors). Since they were so close, we adhered to the Occam's Razor and went with the simplest, six-variable model. This was the model favored by BIC, which gives the greatest penalty for complexity in models. The six variables used in the regression context are then `volatile.acidity`, `chlorides`, `total.sulfur.dioxide`, `pH`, `sulphates`, and `alcohol`. All of the variables from the classification selection appear here with the exception of density. In the regression context, `pH` was the variable that proved to be a rather unintuitive inclusion, since it is similar to density in the sense that it is not easily detected by the five human senses. The inclusion of these two variables in the model portray the importance of the chemical properties of a bottle of wine when it comes to quality.

```{r form}
# Write formula for regression models
form.reg <- quality ~ volatile.acidity + chlorides + pH + 
            total.sulfur.dioxide + sulphates + alcohol

# Write formula for classification models
form.class <- class ~ alcohol + sulphates + density +
              volatile.acidity + total.sulfur.dioxide
```

@@

A visualization of the classification tree can be found below.

```{r tree, fig.align="center", out.height='4in', out.width='4in'}
#### REGRESSION ####
# Fit basic regression tree model
#tree.reg <- train(form.reg, data = trn, method = "rpart",
#                  trControl = trctrl)

# Obtain RMSE and results
#tree.reg.pred <- predict(tree.reg, newdata = tst)
#tree.reg.rmse <- sqrt(mean((tree.reg.pred - tst$quality)^2))

#### CLASSIFICATION ####
# Fit classification tree model
tree.class <- train(form.class, data = trn, method = "rpart",
                    trControl = trctrl)

@@

# Print out classification tree visualization
tree.class.out <- rpart(form.class, data = trn, cp = 0.01971326)
rpart.plot(tree.class.out, main="Classification Tree")
```

@@

```{r rf}
#### REGRESSION ####
# Fit the regression random forest model 
rf.reg <- train(form.reg, data = trn, method = "rf",
                trControl = trctrl, importance = TRUE)

@@

To fit the LASSO regression model, we used the `glmnet` package in R.
```{r lasso}
#### REGRESSION ####
# Fit LASSO to the data
lasso.reg <- train(form.reg, data = trn, method = "glmnet", trControl = trctrl, 
                   tuneGrid = expand.grid(.alpha=1, .lambda=0.00175))

@@

```{r boost}
#### REGRESSION ####
# Fit the regression boosted tree model 
boost.reg <- train(form.reg, data = trn, method = "gbm",
                trControl = trctrl, verbose = FALSE)

@@

```{r mlr}
#### REGRESSION ####
lm.reg <- train(form.reg, data = trn, method = "lm",
                trControl = trctrl)

# Obtain RMSE and results for linear model
lm.reg.pred <- predict(lm.reg, newdata = tst)
lm.reg.rmse <- sqrt(mean((lm.reg.pred - tst$quality)^2))
#lm.reg.rmse
```

@@

```{r create.df}
# Create data frame output for regression models
df.reg <- rbind.data.frame(boost.reg.rmse, rf.reg.rmse,
                           lm.reg.rmse, knn.reg.rmse, 
                           lasso.reg.rmse) %>% round(4)
df.reg$method <- reg
colnames(df.reg) <- c(LETTERS[1:2])
df.reg <- df.reg %>% select("B", "A")
colnames(df.reg) <- c("Regression Method", "Root MSE")

# Create data frame output for classification models
df.class <- rbind.data.frame(tree.class.res, rf.class.res, 
                             svm.class.res, knn.class.res, 
                             glm.class.res) %>% round(4)
df.class$method <- cl
colnames(df.class) <- c("A", "B", "C", "D")
df.class <- df.class %>% select("D", "A", "B", "C")
colnames(df.class) <- c("Classification Method", "Accuracy", "Sensitivity", "Specificity")
```

```{r reg.out}
# Smash tables next to each other using kable
kable(df.reg, caption = "Regression Model Results", format = "latex") %>% 
  row_spec(0, bold=TRUE) %>% collapse_rows(columns = 1, latex_hline = "none") %>% kable_styling(latex_options = "HOLD_position")
```

```{r class.out}
kable(df.class, caption = "Classification Model Results", format = "latex") %>% 
  row_spec(0, bold=TRUE) %>% collapse_rows(columns = 1, latex_hline = "none") %>%
  kable_styling(latex_options = "HOLD_position")
```
